Large-Scale Content-Based Matching of MIDI and Audio Files

### 2015-04-30 paper.tex
Add DTW cost section

(Listening to alignments to find a threshold)

>>>>

\subsection{DTW cost as confidence score}

The ``cost'' of a DTW path $p, q$ through $S$ is calculated by the sum of the distances between the aligned entries of each sequence, or
$$
\sum_{i = 1}^L S[p[i], q[i]]
$$
As-is, the DTW cost is not an appropriate confidence score for alignment/match quality in our application primarily for two reasons.
First of all, the path length can vary greatly across MIDI/audio file pairs depending on $N$ and $M$.
We therefore compute the \textit{mean} distance between aligned constant-Q spectra across the path instead of the sum of distances:
$$
\frac{1}{L}\sum_{i = 1}^L S[p[i], q[i]]
$$
Secondly, the scale of $S$ (and therefore the DTW cost) depends on how similar the synthesized audio is to the ``real'' audio recording, which is not an important factor in the alignment quality.
This can be mitigated by normalizing the DTW cost by the mean value of the submatrix of $S$ through which the DTW past travels:
$$
\mathcal{N} = \sum_{i = \min(p)}^{\max(p)} \sum_{j = \min(q)}^{\max(q)} S[i, j] 
$$
These normalizations were also discussed in \cite{hu2003polyphonic}.
Combining the above, we get our modified DTW cost $c$:
$$
c = \frac{1}{L \mathcal{N}} \sum_{i = 1}^L S[p[i], q[i]]
$$
In practice, we found these two normalization steps to be of critical importance for obtaining a reliable confidence score.

In order to use our normalized DTW cost to determine when a match and alignment was successful, we need a threshold above which the DTW cost indicates a failure.
To estimate this threshold, we manually auditioned 125 alignments and noted whether each was successful, where ``success'' requires that the audio and MIDI were well-synchronized for their entire duration.
From this, we obtained an AUC-ROC score of $0.986$, indicating a highly reliable confidence metric.
As a threshold, we chose $0.78$, above which no alignments where labeled as successful.
